---
title: "Sexual dimorphism in *Fringilla coelebs*"
author: "Emma Rand"
output:
  bookdown::html_document2: default
  bookdown::word_document2: default
  bookdown::pdf_document2: default
bibliography: refs/reference.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                     warning = FALSE,
                     message = FALSE,
                     fig.retina = 3)
options(digits = 2)
```

```{r packages}
source("R/00-pkg.R")
```

# Introduction {-#intro}

The chaffinch, *Fringilla coelebs*, is one of the most widespread and common passerine bird species [@tidyverse][@cramp_1994]. It is sexually dimorphic in plumage with males being brightly coloured with a blue-grey cap and rust-red underparts and females being grey-brown with paler underparts (see Figure \@ref(fig:chaff-photo-fig)). This study aimed to determine whether *F.coelebs* were also sexually dimorphic in size.

<!-- figure legend -->

(ref:chaff-photo-fig) Adult *Fringilla coelebs* Left, male (by Andreas Trepte, CC BY-SA 2.5 <https://creativecommons.org/licenses/by-sa/2.5>, via Wikimedia Commons); right, female (4028mdk09, CC BY-SA 3.0 <https://creativecommons.org/licenses/by-sa/3.0>, via Wikimedia Commons)

<!-- end of fig legend -->

```{r chaff-photo-fig, out.width="40%", fig.show='hold', fig.cap='(ref:chaff-photo-fig)'}
pics <- c("images/Chaffinch_(Fringilla_coelebs)male.jpg",
          "images/Fulda_Buchfinkweibchen_Juni_2012female.JPG")
knitr::include_graphics(pics)
```

# Methods {-#methods}
```{r import}
# chaff_tidy
source("R/01-import.R")
```

```{r summarise}
# outputs chaff_summarymary with columns mean, std, n, se, ssx
source("R/02-summarise.R")
```


```{r extract-summary}
# extract values for inline reporting
# ss(x)
m_ss <- chaff_summary$ssx[chaff_summary$sex == "males"]
f_ss <- chaff_summary$ssx[chaff_summary$sex == "females"]
# sample sizes
n_males <- chaff_summary$n[chaff_summary$sex == "males"] 
n_females <- chaff_summary$n[chaff_summary$sex == "females"] 
# highest mean, se and the sex to which they belongs
highestmean <- max(chaff_summary$mean) 
highestse <- chaff_summary$se[chaff_summary$mean == highestmean] 
highestsex <- chaff_summary$sex[chaff_summary$mean == highestmean]
# lowest mean, se and the sex to which they belongs
lowestmean <- min(chaff_summary$mean) 
lowestse <- chaff_summary$se[chaff_summary$mean == lowestmean] 
lowestsex <- chaff_summary$sex[chaff_summary$mean == lowestmean]
```

We randomly sampled `r n_males` male and `r n_females` female birds on the stray and determined their mass with spring scales. Analysis was carried out with R version 4.1.1 [@R-core] and tidyverse packages [@tidyverse].


# Results {-#results}

```{r testing}
# outputs mod, a t-test object
source("R/03-test.R")
```

```{r extract-testing}
res <- summary(mod)
tval <- res$coefficients["sexmales", "t value"]
df <- res$df[2]
if (res$coefficients["sexmales", "Pr(>|t|)"] < 0.001) {
        b1p = "< 0.001"
        }
if (res$coefficients["sexmales", "Pr(>|t|)"] > 0.001) {
        b1p = paste("=", round(res$coefficients["sexmales", "Pr(>|t|)"], 3))
        }
rsq <- res$r.squared
```

We wouldn't normally report the sum of the squared deviations ($SS(x)$) but we do so here to make use of a function. For males, the $SS(x)$ = `r m_ss` and for females, the $SS(x)$ = `r f_ss`


`r str_to_sentence(highestsex)` ($\bar{x} \pm s.e.$: `r highestmean` $\pm$ `r highestse` g) have significantly higher mass than `r lowestsex` (`r lowestmean` $\pm$ `r lowestse` g)($t$ = `r tval`; $d.f.$ = `r df`; $p$ `r b1p`). See figure \@ref(fig:chaff-fig).

```{r chaff-fig, fig.height=4, fig.width=4, fig.cap="(ref:chaff-fig)"}
source("R/04-plot.R")
fig1
```

<!-- figure legend -->

(ref:chaff-fig) Mass of adult chaffinches caught on the Stray. Error bars are $\bar{x} \pm 1 s.e.$  

<!-- end of fig legend -->

I've also gratuitously included a table with the same information just for the sake of including a table. See Table \@ref(tab:summary-table).

```{r summary-table}
knitr::kable(chaff_summary,
             digits = 2,
             caption = 'A summary of the data.',
             row.names = FALSE)
```

# Discussion {-#discussion}

Here we pick up points from the introduction. 

## Word count {-}
Word count calculated with package **`wordcountaddin`** [@wc]
```{r wordcount}
wc_main <- wordcountaddin:::text_stats("main.Rmd")[3] %>% str_extract("[0-9]+") %>% as.numeric()
wc_readme <- wordcountaddin:::text_stats("README.md")[3] %>% str_extract("[0-9]+") %>% as.numeric()
```

This doc: `r wc_main`  
The README: `r wc_readme`  
Total: `r wc_main + wc_readme`  

# References {-}
